Software architectures to integrate workflow engines in science gateways

### 2016-10-10 answers.tex
Started answering comments from reviewer 1

\answer{}

@@

\answer{}

@@

@@

\begin{review}
 Given these observations and the other measurements provided by the paper, the overall performance presented in Table 3, where the Workflow Conversion outperforms the Service Invocation pattern, is considered counter intuitive. The result is justified though by the evaluation of the Service-invocation pattern. Here, a high value (low score) is assigned to the Extensibility metrics E1, concerning the inclusion
of new workflow engines. This seems to be too negative, since a service invocation approach can rely on generic interfaces, re-purposing via a common service layer supporting multiple technologies. This is reflected also in the negative evaluation of the S1,S2 metrics for the same pattern. On the contrary, the choice of this architecture would intuitively suggest re-usable, scalable and decoupled solutions. Eventually, the arguments presented were not considered convincing.
\end{review}

\begin{review}
Again on the definition of Robustness. It introduces two metrics: Specific Components and Interactions. These are explained in terms of what they not include but no evidence or examples are provided to better understand what is intended specific. Moreover, some interactions which are mentioned as being not specific, become specific during the evaluation. For instance, interaction 'c', which is defined as the upload of input data and results (c1) and to transfer processed data across infrastructures (c2), are initially considered as an example of not specific interaction. Though, when measuring the Robustness of the Pool Model,  c2 suddenly appears in the count,  Does the definition of R2 means that non-specific interactions should be counted? The higher number the worse robustness?
\end{review}

>>>>

\answer{We thank the reviewer for the thorough review and useful comments.

  The results and discussion reported in our paper are indeed aimed at
  an expert audience of science gateway and workflow engine
  developers. The evaluation metrics have been designed based on our
  long experience with both science gateways and workflow engines, and
  we completely acknowledge that non experts may misinterpret the
  results.

  The fact that the patterns may be mixed in real implementations is,
  in our opinion, another motivation for our abstraction effort as it
  helps ``untangling'' the various architectural concepts that may be
  present in a real system. This is particularly true for the abstract
  architectures (nested workflows and workflow conversion) that may be
  instantiated in various ways. Our abstractions highlight that other
  instantiation models may be worth designing, for instance nested
  workflows with task encapsulation, as mentioned in Section 4.2. in
  the Discussion.

  We added the following paragraph to the Introduction to reflect this discussion:}

\revised{The analysis presented in the paper is intended for
experts in science gateway and workflow engine design. It is an
abstraction effort to identify and evaluate the fundamental
architectural patterns that are encountered while integrating workflow
engines and science gateways. In real systems, such patterns are often
intertwined due to considerations related to the historical and
technical context of software projects. The following work helps
identify the key patterns in systems and highlight potential
improvements.}

@@

\answer{The actual workflow engine deployed in the infrastructure
  requires the gateway to directly control its own sub-tasks. We
  revised the wording of Section 2.4 to clarify this point.}

@@

\answer{S1 denotes the ability of the system to start multiple engine
  instances to execute different workflows, while S2 denotes the
  possibility to execute a single workflow with multiple engines. We
  revised the wording in the description of S1 (page 11) to clarify
  this point. We find the current description of S2 clear enough on
  this point: ``[S2] measures the ability to distribute the execution of
  \emph{a single workflow} among different engine instances.''}

@@

\answer{We introduced E3 precisely to account for the additional
  complexity introduced by interaction (g) when adding new
  workflows.

It should be understood here that workflow conversion is
  presented as a pattern to integrate workflow \emph{engines} in a
  science gateway, through workflow translation to a language already
  supported by the gateway. We believe that integrating new workflow
  engines (therefore new workflows) through workflow conversion is a
  strong architectural design choice that may, depending on the
  availability of a reliable interaction (g), be avoided by
  integrating workflow engines using a different pattern. We added the following sentence at the beginning of Section 2.7 to explain that:

\revised{Workflow conversion is presented here as a pattern to
integrate workflow \emph{engines} in a science gateway, through
workflow format conversion from a native format to the science gateway
format.}
}

We definitely agree with the reviewer that interaction (g) is
  a tricky one that may lead to critical errors. We added the
  following sentence to the paragraph on interaction (g) in Section
  4.2. of the discussion:

\revised{If not properly validated, workflow language conversion
  could introduce critical errors impacting the robustness and
  correctness of the execution. }

This sentence complements the note already made in section 3.7:
``Note, however,
that implementing interaction g can require sub-
stantial work depending on the complexity of the
workflow language used by the new engine.``

\begin{review}
 Given these observations and the other measurements provided by the paper, the overall performance presented in Table 3, where the Workflow Conversion outperforms the Service Invocation pattern, is considered counter intuitive. The result is justified though by the evaluation of the Service-invocation pattern.
\end{review}

\answer{As discussed previously, we share the reviewer's concerns
  about the feasibility of workflow conversion (see comments on
  interaction (g) in Section 4.2). The performance of workflow
  conversion is strongly related to the availability of a validated
  interaction (g), which is very dependent on the particular workflow
  engines under consideration. We believe that the issue raised by the
  reviewer basically comes down to the weight given to software
  components and interactions in the analysis. However, using
  different weights for different components or interactions would
  open an infinite number of configurations for discussion. For
  instance, in some cases, the workflow engine already comes with a
  workflow service which therefore doesn't have to be included. To
  address this issue, we shared a spreadsheet at
  \url{https://frama.link/LrPRsCQp} that allows editing the following
  weights:
  \begin{itemize}
  \item Weight given to software components and interactions in
  ``integration'', ``extensibility'' and ``robustness'' criteria.
  \item Weight given to metrics in a criterion (e.g., E$_1$ vs. E$_2$ in
  ``extensibility'').
  \item Weight given to criteria in overall evaluation of Table 3.
\end{itemize}
The paper references this spreadsheet wherever equal weights are
assumed. The readers and reviewers are invited to rely on this
spreadsheet for any issue related to the importance given to a
specific component, interaction, metric or criterion in the
analysis. In the paper, we assumed that all weights equal 1, which we
believe is a reasonable but limited choice.
}

\begin{review}
  Here, a high value (low score) is assigned to the Extensibility
  metrics E1, concerning the inclusion of new workflow engines. This
  seems to be too negative, since a service invocation approach can
  rely on generic interfaces, re-purposing via a common service layer
  supporting multiple technologies.
\end{review}

\answer{The reviewer is right that re-purposing software components
  and interactions can lead to revisit the evaluation. We updated the following sentence in section 4.2. to better acknowledge this point:}

The particular technical or historical context of a science gateway
project may also influence the choice of an architecture to
integrate workflow engines. For instance, many workflow engines are
already available as web services, which tends to favor service
invocation (\revised{in particular when interactions can be re-purposed}), and other science gateways may have strongly tested task
and data control features (interactions \texttt{b} and \texttt{c}),
which would favor task encapsulation. Similarly, adding a new type of engine
may facilitate integration of a new infrastructure when interactions
\texttt{b} and \texttt{c$_2$} are already available. The migration
cost between architectures has been ignored as well.

In the specific case of service invocation, we agree with the reviewer
that, when adding a new workflow engine, the workflow service or interaction \texttt{d} can always be repurposed depending on the context. We decremented E1 and added the following explanation in section 3.3.:

\revised{However, when a new type of workflow engine is added, it is always possible to reuse either an existing available workflow service or interaction \texttt{d} (when a new workflow service has to be developed). Therefore, \texttt{E$_1$}=3}.

\begin{review}
 This is reflected also in the negative evaluation of the S1,S2 metrics for the same pattern. On the contrary, the choice of this architecture would intuitively suggest re-usable, scalable and decoupled solutions.
\end{review}

\answer{While we agree with the reviewer that re-usability should
  definitely be taken into account in E$_1$, we do not see how it
  could help S$_1$ and S$_2$.

  For S$_1$, starting new instances of the workflow service still
  requires that a specific mechanism is implemented in the science
  gateway to distribute the load among these services, even if all
  these services share the same interface. Again, the difficulty to
  implement such mechanisms could be extensively discussed: some
  systems may rely on a simple DNS-level load balancing, but this
  would not cope with the fact that workflow executions have different
  resource requirements and that workflow services may have different
  availabilities (for instance, a workflow service may crash in case a
  workflow execution filled up its disk); other systems may balance
  the load based on resource usage (e.g. RAM, CPU and disk) in the
  workflow services, but this requires an elaborated monitoring
  infrastructure. On the contrary, adding new engine instances is
  naturally supported by task encapsulation and pool, which justifies
  that S$_1=0$ for these models while S$_1$=1 for service invocation.

  For S$_2$, we cannot think of a practical reason why the service
  invocation model would help distribute a single workflow execution
  in multiple engine instances. In our opinion, this feature is
  exclusive to the nested workflow model. }

\begin{review}
 Eventually, the arguments presented were not considered convincing.
\end{review}

We hope that the update made in the extensibility scores and the
explanations given for scalability convinced the reviewer.

\begin{review}
  Again on the definition of Robustness. It introduces two metrics:
  Specific Components and Interactions. These are explained in terms
  of what they not include but no evidence or examples are provided to
  better understand what is intended specific. 
\end{review}

\answer{We updated the definition of R$_1$ and R$_2$ to include
  positive examples as well as negative ones. In addition, the
  specific components included in these metrics are represented in red
  on Figure 2.}

\begin{review}
Moreover, some
  interactions which are mentioned as being not specific, become
  specific during the evaluation. For instance, interaction 'c', which
  is defined as the upload of input data and results (c1) and to
  transfer processed data across infrastructures (c2), are initially
  considered as an example of not specific interaction. Though, when
  measuring the Robustness of the Pool Model, c2 suddenly appears in
  the count, 
\end{review}

\texttt{c}, \texttt{c$_1$} and \texttt{c$_2$} are three different
interactions, which explains the confusion. \texttt{c$_2$} is always
specific to workflow execution because it is defined as the
interaction between the workflow engine and the infrastructure, used
to transfer workflow data (inputs, outputs or intermediate
results). \texttt{c$_1$} is not specific to workflow execution because
it is used by the science gateway to transfer any data regardless of
workflow executions. Letter \texttt{c} was used on the task
encapsulation diagram in Figure 2 to represent that the science gateway transfers
both user and workflow data. We now use \texttt{c$_1$} for all the
data transfers initiated by the science gateway, and \texttt{c$_2$}
for all the data transfers initiated by the workflow engine,
regardless of the type of data transferred. Consequently, we now use
\texttt{c$_1$} instead of \texttt{c} on the task encapsulation diagram in Figure 2.

The ambiguity is now removed and the new definition of R$_1$ and R$_2$ reads as follows:

Specific components (\texttt{R$_1$}): number of specific components involved in workflow executions. \revised{For instance, the
  workflow service is a specific component because its main objective
  is to execute workflows. On the contrary, the science gateway is
  \emph{not} specific to workflow execution since it is used for a
  variety of other functions such as user authentication, data
  transfer, etc.}

Specific interactions (\texttt{R$_2$}): number of specific
  interactions involved in workflow executions. \revised{For
    instance, the pool-agent interaction (interaction \texttt{f}) is
    specific to workflow execution while data control (interaction
    \texttt{c$_1$}) is \emph{not} because it is used by the science gateway to transfer
    user data regardless of a particular workflow execution}.

\begin{review}
Does the definition of R2 means that non-specific
  interactions should be counted?
\end{review}

\answer{No it doesn't. We hope that the new examples in the definition of
R$_1$ and R$_2$ and the removal of any reference to interaction
\texttt{c} clarifies this point.}

\begin{review}
The higher number the worse
  robustness?
\end{review}

\answer{Yes, as for any other metric, high values indicate worse
performance. This is explained at the beginning of section 3:
``Criteria break down into specific met- rics where lower value
indicates better performance.''}
